Fractional land cover classification method assessment using PROBA-V satellite data

### 2019-08-06 article/article.tex
Changes from papeeria

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS}{CGLS-LC100}{Copernicus Global Land Services Land Cover}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym[text={random forest}]{RF}{RF}{random forest}
\makenoidxglossaries
% Disable abbreviating RF for now
\glsunset{RF}

@@

In addition to classification methods, one aspect that is important for classification accuracy is the data (covariates) that the algorithm is trained on and predict land cover from \citep{yu2014metadiscoveries}. In addition to spectral covariates (blue, red, near infrared, shortwave infrared bands), temporal covariates (growing season start, duration, growing intensity) are commonly used to improve the separability of crops \citep{jakubauskas2001harmonic}. Elevation covariates (elevation, slope, aspect) are known to improve the separability between tree classes \citep{burrough2001fuzzy}. Vegetation indices have been used to improve the detection of wetlands \citep{sader1995wetlands}, as well as for defining classification rules in rule-based fuzzy classification \citep{baraldi2006rulebased}. Therefore in this thesis such covariates have also been used in order to increase land cover classification accuracy, and their importance for prediction was analysed.

@@

\subsubsection{Optical covariates}

@@

\begin{enumerate}
 \item GLSDEM
 \item Proba-V time series
 \item SoilGrids and LandGIS
 \item WorldClim
\end{enumerate}

\subsection{Preprocessing}

>>>>

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS}{CGLS-LC100}{Copernicus Global Land Services Land Cover}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym[text={random forest}]{RF}{RF}{random forest}
\makenoidxglossaries
% Disable abbreviating RF for now
\glsunset{RF}

@@

In addition to classification methods, one aspect that is important for classification accuracy is the data (covariates) that the algorithm is trained on and predict land cover from \citep{yu2014metadiscoveries}. In addition to spectral covariates (blue, red, near infrared, shortwave infrared bands), temporal covariates (growing season start, duration, growing intensity) are commonly used to improve the separability of crops \citep{jakubauskas2001harmonic}. Elevation covariates (elevation, slope, aspect) are known to improve the separability between tree classes \citep{burrough2001fuzzy}. \Glspl{VI} have been used to improve the detection of wetlands \citep{sader1995wetlands}, as well as for defining classification rules in rule-based fuzzy classification \citep{baraldi2006rulebased}. Therefore in this thesis such covariates have also been used in order to increase land cover classification accuracy, and their importance for prediction was analysed.

@@

%\begin{enumerate}
% \item GLSDEM
% \item Proba-V time series
% \item SoilGrids and LandGIS
% \item WorldClim
%\end{enumerate}

\subsubsection{Optical covariates}
% * <Dainius Masiliūnas> 16:04:43 06 Aug 2019 UTC+0200:
% Maybe better to call it "Proba-V data" or such

The entire archive of the Proba-V 100 m Level 3 Top-of-Canopy 5-day composite product was used for this study.
The time series of NDVI and each of the bands at the points of interest were cloud masked, first by applying the status mask provided with the product itself, and then by running a temporal cloud filter to remove the remaining outliers that were too far from the fitted \ac{LOESS} curve over the blue band.

The resulting cloud-free time series was then used to extract descriptive statistics (mean and interquartile range) of NDVI over the whole time series.
Next, harmonic analysis was run in order to decompose the time series into sine and cosine components for two frequencies (annual and semiannual), as well as the trend and intercept of the model.
From this, phase and amplitude for the two frequencies were calculated.

Lastly, extra \glspl{VI} (NDMI, EVI, OSAVI and NIR\textsubscript{v}) were calculated.
Each of these \glspl{VI} was then composited into a single image that represents the leaf-on season of 2015.

The optical preprocessing chain is displayed graphically in figure \ref{fig-preprocessing-optical}.
All of the outputs were then used in the processing chain.

@@

\subsubsection{Terrain covariates}

The NASA \ac{GLSDEM} 90 m product was used as the elevation covariate.
In addition, terrain parameters were calculated out of elevation: slope, aspect, roughness and \ac{TPI}.
All terrain covariates were then resampled to the Proba-V grid.

\subsubsection{Climate covariates}

The climate data from WorldClim 2.0 30 s product was used.
Biophysical parameters were calculated from the data as well.
Some additional biophysical parameters were also calculated, namely the means, minimums and maximums over the whole time series of solar radiation, wind speed and water vapour pressure.

\subsubsection{Soil covariates}

\subsubsection{Location covariates}

The latitude, longitude and absolute latitude of the points were also included as covariates.

\subsection{Covariate selection}
